Security vs. Maintainability: Fixing Vulnerabilities Obfuscates your Code

### 2019-01-27 paper/ICPC19/icpc19.tex
section 2 finished

@@

@@

One example is the refactoring of the DTLS protocol implementation in OpenSSL after
Adam Langley detect that remote attackers could cause a Denial-of-Service (DoS) attack
when crafting DTLS handshake messages to trigger memory allocations corresponding to large length values. 
This vulnerability is listed at the Common Vulnerabilities and Exposures dictionary as CVE-2014-3506\footnote{CVE-2014-3506 details available at http://cve.mitre.org/cgi-bin/cvename.cgi?name=CVE-2014-3506 (Accessed on January 25, 2019)} and 
it is one of the vulnerabilities under evaluation in this study. The impact of this refactoring
on the OpenSSL maintainability was negative. The following snippet presents the codebase changes 
performend on the \emph{ssl/d1\_both.c} file\footnote{CVE-2014-3506 fix available  at https://github.com/openssl/openssl/commit/\\1250f12613b61758675848f6600ebd914ccd7636} by the OpenSLL developers to fix the vulnerability.

@@

In the \emph{dtls1\_reassemble\_fragment} function the value of \emph{msg\_hdr-$>$frag\_off+frag\_len} was checked against the
maximum hadshake message size but \emph{msg\_len} is allocated to the fragment buffer in line 37. Thus, if the fragment was within
the allowed size, the pending handshake message could consume $16MB$+$2MB$. According with the message commit, 10 handshake messages
are allowed, so the attacker could consume approximately $180MB$ per DTLS connection.

@@

\ref{lst:func1} $dtls1\_max\_handshake\_message\_len$ function was inserted to limit the number of bytes allowed in a DTLS handshake message for \emph{s}. This function replaces the lines 18-21.

\ref{lst:func2} Checks if the maximum certificate list size (\emph{s-$>$max\_cert\_list}) requires more length than the maximum length (\emph{max\_len}). If yes, it return the certificate list size, if not it return the maximum length.

@@

Although limiting the size of a handshake message seems an easy problem to fix, it is possible to see that there was a considerable amount of changes performed
in the codebase which led to a decrease in the project maintainability.

@@

@@

\begin{framed}
\textit{\textbf{RQ3} }
\end{framed}

\section{Methodology}

@@

In total, the dataset contains 716 security refactorings. These refactorings were computed by the BCH tool to calculate maintainability. However, some of the refactorings are not considered in the final results due to limitations of BCH: lack of language support and project size. The Wilcoxon statistical test - described more in-depth in subsection \ref{sec:statsval} - is performed to evaluate the significance of our results. Each type of security flaw is tested, Figure \ref{fig:pat}. One of the main requirements of this test is that the input sample needs to comprise more than 20 commits. Thus, the patterns not complying with this requirement were reallocated to the \textit{Miscellaneous} pattern. In the end, we highlight the following patterns:

@@

\textcolor{mypink3}{@todo: Discuss pratt improvement and accuracy}

\section{Results}

@@

@@

@@

\section{Discussion}

\section{Threats to Validity}

\section{Related Work}

\section{Conclusion and Future Work}

\section*{Acknowledgment}

Better Code Hub?

>>>>

Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.

@@

This paper is structured as follows: section \ref{sec:motivation} introduces an example of a security refactoring of a known vulnerability found in OpenSSL\footnote{Repository details available at https://github.com/openssl/openssl (Accessed on January 27, 2019)} software; section \ref{sec:methodology} describes the methodology used to answer the research questions; section \ref{sec:results} presents the results and section \ref{sec:discussion} discusses their implications; section \ref{sec:threats}, threats to the validity of the study are enumerated; section \ref{sec:rw} describres the different work and existing literature in the field of study; and, finally section \ref{sec:conclusions} concludes the main findings and elaborates on future work.

@@

One example is the refactoring of the DTLS protocol implementation in OpenSSL after
Adam Langley detect that remote attackers could cause a Denial-of-Service (DoS) attack
when crafting DTLS handshake messages to trigger memory allocations corresponding to large length values. 
This vulnerability is listed at the Common Vulnerabilities and Exposures dictionary as CVE-2014-3506\footnote{CVE-2014-3506 details available at http://cve.mitre.org/cgi-bin/cvename.cgi?name=CVE-2014-3506 (Accessed on January 27, 2019)} and 
it is one of the vulnerabilities under evaluation in this study. The impact of this refactoring
on the OpenSSL maintainability was negative. The following snippet presents the codebase changes 
performend on the \emph{ssl/d1\_both.c} file\footnote{CVE-2014-3506 fix available  at https://github.com/openssl/openssl/commit/\\1250f12613b61758675848f6600ebd914ccd7636 (Accessed on January 27, 2019)} by the 
OpenSLL developers to fix the vulnerability.

@@

In the \texttt{dtls1\_reassemble\_fragment} function the value of \texttt{msg\_hdr-$>$frag\_off+frag\_len} was checked against the
maximum hadshake message size but \texttt{msg\_len} is allocated to the fragment buffer in line 37. Thus, if the fragment was within
the allowed size, the pending handshake message could consume $16MB$+$2MB$. According with the message commit, 10 handshake messages
are allowed, so the attacker could consume approximately $180MB$ per DTLS connection.

@@

\ref{lst:func1} \texttt{dtls1\_max\_handshake\_message\_len} function was inserted to limit the number of bytes allowed in a DTLS handshake message for \texttt{s}. This function replaces the lines 18-21.

\ref{lst:func2} Checks if the maximum certificate list size (\texttt{s-$>$max\_cert\_list}) requires more length than the maximum length (\texttt{max\_len}). If yes, it return the certificate list size, if not it return the maximum length.

@@

Although limiting the size of a handshake message seems an elementary problem to repair, the example shows that there was a significant amount of changes performed in the codebase which produced a negative impact in the project maintainability. Even though the developer tries to simplify the codebase with the creation of a smaller unit \texttt{dtls1\_max\_handshake\_message\_len}, it also disrupts 2 of the guidelines proposed by the Software Improvement Group (SIG) for building maintainable software~\cite{Visser:2016:OREILLY}. It adds a new branch point to the \texttt{dtls1\_process\_out\_of\_seq\_message} unit (\ref{lst:func5}) increasing the cyclomatic complexity and breaking the \emph{Write Simple Units of Code} guideline. And, it increases the codebase size in 4 LOCs (13 lines added, 9 lines deleted) which infringes the \emph{Keep Your Codebase Small} guideline.

In this study, our concern is if while improving software security, developers are also inserting a negative impact on the maintainability of their software which may increase technical debt when choosing the easiest solution instead of the using the best approach. To answer the following two research questions, we use a dataset of security refactorings to measure the impact of the changes in the maintainability of open-source software.

@@

Often security flaws require refactoring code to make software more secure. However, there is no evidence yet of how security refactorings impact the maintainability of open-source software. Our suspicion is that developers tend to introduce technical debt in their software when refactoring the codebose to address a security flaw because they tend to choose the easiest path to solve it. To address it, we compute the maintainability value of $607$ commits using \emph{Better Code Hub}. The same approach is applied to a randomly generated dataset of regular commits - baseline - that we use to understand how maintainability evolves when security refactorings are performed versus when they are not.

@@

There are security flaws that are more difficult to refactor than others. For instance, introducing a mechanism to secure authentication is not so easy to implement as fixing a cross-site scripting vulnerability with a \texttt{escape} function. Knowing which patterns are more likely to increase maintainability issues is one step forward to bring awareness to security engineers of what patterns need more attention. The taxonomy of security patterns used to answer this question is the same as the one provided by the authors of the dataset. Maintainability is measured separately for each pattern.

\section{Methodology}\label{sec:methodology}

@@

In total, the dataset contains 716 security refactorings. These refactorings were computed by the BCH tool to calculate maintainability. However, some of the refactorings are not considered in the final results due to limitations of BCH: lack of language support and project size. The final dataset contains $607$ security refactorings. The Wilcoxon statistical test - described more in-depth in subsection \ref{sec:statsval} - is performed to evaluate the significance of our results. Each type of security flaw is tested, Figure \ref{fig:pat}. One of the main requirements of this test is that the input sample needs to comprise more than 20 commits. Thus, the patterns not complying with this requirement were reallocated to the \textit{Miscellaneous} pattern. In the end, we highlight the following patterns:

@@

\textcolor{mypink3}{@todo: Discuss pratt improvement and accuracy} Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged.

\section{Results}\label{sec:results}

Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged.

Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged.

Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged.

@@

\section{Discussion}\label{sec:discussion}

@@

Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged.

@@

Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged.

\section{Threats to Validity}\label{sec:threats}

Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book.

\section{Related Work}\label{sec:rw}

Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged.

\section{Conclusion and Future Work}\label{sec:conclusions}

It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged. It was popularised in the 1960s with the release of Letraset sheets containing Lorem Ipsum passages, and more recently with desktop publishing software like Aldus PageMaker including versions of Lorem Ipsum.Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged.

\section*{Acknowledgment}\label{sec:ack}

We thank to the Software Improvement Group (SIG) team for all the support and help in validating our methodology.
